---
title: "Mixed Models Inventory"
author: "Johnny van Doorn"
runtime: shiny
output: html_document
---
<style type="text/css">
.main-container {
max-width: 1600px;
margin-left: auto;
margin-right: auto;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(BayesFactor)
library(lme4)
library(lmerTest)
library(afex)
source('mixedEffectFunctions.R')

```

# Table of contents
1. [Defining the Models](#defining)
2. [Shiny App](#shiny)




## Defining the Models <a name="defining"></a>
Suppose $I$ people each observe $M$ trials in $J$ conditions. We have the following candidate models, each with different implications.

Model 1, Weak Null, intercept ($\mu$) only:
\[
Y_{ijm} \sim \mbox{N}(\mu,\sigma^2)
\]


Model 2,  No subject effect, fixed effect ($\nu$) of condition only ($x_j=-1/2,1/2$ if $J = 2$):
\[
Y_{ijm} \sim \mbox{N}(\mu+x_j\nu,\sigma^2)
\]

Model 3, Strong Null, no condition effects, but random intercepts per subject (i.e., random intercept per subject):
\[
\begin{aligned}
Y_{ijm} &\sim \mbox{N}(\alpha_i,\sigma^2)\\
\alpha_i &\sim \mbox{N}(\mu,\delta_\alpha)
\end{aligned}
\]



Model 4, No subject-by-treatment interactions, random effect of subject and main effect of condition:
\[
\begin{aligned}
Y_{ijm} &\sim \mbox{N}(\alpha_i+x_j\nu,\sigma^2)\\
\alpha_i &\sim \mbox{N}(0,\delta_\alpha)
\end{aligned}
\]


Model 5, Main effect with subject-by-treatment interactions (i.e., random slopes per subject):
\[
\begin{aligned}
Y_{ijm} &\sim \mbox{N}(\alpha_i+x_j(\nu+\theta_i),\sigma^2)\\
\theta_i &\sim \mbox{N}(0,\delta_\theta)\\
\alpha_i &\sim \mbox{N}(\mu,\delta_\alpha)
\end{aligned}
\]

## Shiny App  <a name="shiny"></a>

```{r, echo = FALSE}
server <- function(input, output) {
  mydat <- reactive({
    set.seed(input$seed)
    myDataSim(fixedEffectItem = input$fixedEffectItem, randomSlopeItemSigma = input$randSlopeItem,
              randomInterceptItemSigma = input$randInterItem,
              fixedEffectCondition = input$fixedEffectCond, randomSlopeConditionSigma = input$randSlopeCond,
              randomInterceptConditionSigma = input$randInterCond, errorSigma = input$errorSig,
              randomRhoSlopeInterceptCondition = input$rhoEffectsCondition, 
              randomRhoSlopeInterceptItem = input$rhoEffectsItem, 
              errorSigBetween = 0,
              nSub = input$nSub, nItem = input$nItem, nCond = input$nCond)})
  
  output$bfTable <- renderTable({
    gbfs <- generalTestBF(y  ~ cond * id, whichRandom = c('id', 'cond:id'), data = mydat(),
                          whichModels = input$whichModels,
                          rscaleFixed = input$fixedPrior , rscaleRandom = input$randomEffectPrior, 
                          rscaleCont = 1)@bayesFactor
    return(data.frame(model = rownames(gbfs), logBF_intercept = gbfs[, 1], logBF_id = c(NA, 0, gbfs[-(1:2), 1] - gbfs["id",1])))
    # return(data.frame(logBF_intercept = 1:4))#, logBF_id = c(NA, 0, allbf)))
    
  }, caption = "Model comparison full data: Bayes factors with either Model 1 or Model 3 as the null model.", striped = TRUE)
  
  output$bfTableAgg <- renderTable({
    dat <- mydat()
    mydatAgg <- as.data.frame.table(tapply(dat$y,list(dat$id,dat$cond),mean)) # same as mydatUnivar
    colnames(mydatAgg)=c("id","cond", "y")
    
    gbfs <- generalTestBF(y  ~ cond * id, whichRandom = c('id', 'cond:id'), data = mydatAgg, 
                          whichModels = input$whichModels, rscaleFixed = input$fixedPriorAgg , 
                          rscaleRandom = input$randomEffectPriorAgg, 
                          rscaleCont = 1)@bayesFactor
    return(data.frame(model = rownames(gbfs), logBF_intercept = gbfs[, 1], logBF_id = c(NA, 0, gbfs[-(1:2), 1] - gbfs["id",1])))
  }, caption = "Model comparison aggregated data: Bayes factors with either Model 1 or Model 3 as the null model.", striped = TRUE)
  
  output$linesPlot <- renderPlot({
    mylwd <- 2
    dat <- mydat()
    mydatAgg <- as.data.frame.table(tapply(dat$y,list(dat$id,dat$cond),mean)) 
    colnames(mydatAgg)=c("id","cond", "y")
    
    par(mfrow = c(1, 2), cex = 1.2)
    palette(rainbow(input$nSub))
    allCols <- adjustcolor(palette(), alpha.f = 0.4)
    plot(dat$cond, dat$y, bty = "n", las = 1, type ="n", xlab = "Condition", ylab = "y", main = "Full Data")
    
    points(as.numeric(dat$cond), dat$y)
    for (i in 1:input$nSub) {
      ss <- subset(dat, id == i)
      for (j in 1:input$nItem) {
        sss <- subset(ss, item == j)
        lines(1:input$nCond, sss$y, col = allCols[i], lwd = mylwd)
      }
    }
    
    boxplot(dat$y ~ dat$cond, bty = "n", las = 1, type ="n", xlab = "Condition", ylab = "y", main = "Aggregated Data", 
            col = "white", border = "white", bg = "white")
    boxplot(mydatAgg$y ~ mydatAgg$cond, add = TRUE, las = 1)
    points(as.numeric(mydatAgg$cond), mydatAgg$y)
    for (i in 1:input$nSub) {
      ss <- subset(mydatAgg, id == i)
      lines(1:input$nCond, tapply(ss$y, ss$cond, mean), col = allCols[i], lwd = mylwd)
    }
    
  })
  
  output$data <- downloadHandler(
    filename = function() {
      paste('mixedEffectsData_', Sys.Date(), '.csv', sep='')
    },
    content = function(con) {
      write.csv(mydat(), con)
    }
  )
}

ui <- fluidPage (
  
  sidebarLayout(            
    sidebarPanel(
      fluidRow(
        column(10,                
               numericInput('seed', "Set seed", value = 123, min = 1, step = 1),
               downloadButton(outputId = 'data', label = "Download data"),
               radioButtons("whichModels", "Which Models?", choices = c("withmain", "all"), selected = "withmain")
        ),        
        column(6,
               sliderInput('nSub', '#sub', value = 10, min = 2, max = 100, step = 1),
               sliderInput('nCond', '#condition', value = 2, min = 2, max = 10, step = 1),
               sliderInput('fixedEffectCond', 'Fixed effect cond', value = 0, min = 0, max = 3, step = 0.1),
               sliderInput('randSlopeCond', 'Random slope cond sigma', value = 0, min = 0, max = 3, step = 0.1),
               sliderInput('randInterCond', 'Random intercept cond sigma', value = 0, min = 0, max = 3, 
                           step = 0.1),
               sliderInput('rhoEffectsCondition', 'Rho intercept/slope cond', value = 0, min = -1, max = 1, 
                           step = 0.1)
        ),
        column(6,
               sliderInput('errorSig', 'Error sigma', value = 1, min = 0, max = 10, step = 0.1),
               sliderInput('nItem', '#item', value = 2, min = 1, max = 100, step = 1),
               sliderInput('fixedEffectItem', 'Fixed effect item', value = 0, min = 0, max = 3, step = 0.1),
               sliderInput('randSlopeItem', 'Random slope item sigma', value = 0, min = 0, max = 3, step = 0.1),
               sliderInput('randInterItem', 'Random intercept item sigma', value = 0, min = 0, max = 3, 
                           step = 0.1),
               sliderInput('rhoEffectsItem', 'Rho intercept/slope item', value = 0, min = -1, max = 1, 
                           step = 0.1)
        ),
        column(6,
               sliderInput('fixedPrior', 'Prior width fixed effect (full)', value = 0.2, min = 0.1, max = 3, step = 0.1),
               sliderInput('randomEffectPrior', 'Prior random effects (full)', value = 1, min = 0.1, max = 3, step = 0.1),
        ),
        column(6, 
               sliderInput('fixedPriorAgg', 'Prior width fixed effect (agg)', value = 0.2, min = 0.1, max = 3, step = 0.1),
               sliderInput('randomEffectPriorAgg', 'Prior random effects (agg)', value = 1, min = 0.1, max = 3, step = 0.1)
               
        )
      )),
    
    mainPanel(
      plotOutput("linesPlot"),
      fluidRow(
        column(6, tableOutput("bfTable")),
        column(6, tableOutput("bfTableAgg"))),
      
      tags$head(tags$style(HTML("
                               body {
                                  width: 100% !important;
                                  max-width: 120% !important;
                               }

                               ")))
    )
  )
)
shinyApp(ui, server, options = list(width = "100%", height = 1000))
```


